The \project{unRAVE} catalog of X.X million detailed stellar abundances for chemodynamic studies

### 2016-08-16 article/unrave.tex
Minor things

The Milky Way is considered to be the best laboratory for understanding galaxy
formation and evolution.  This premise hinges on the ability to precisely measure 
the position, motion, and chemistry of (many) individual stars, and to infer the 
structure, kinematics, and chemical enrichment of the Galaxy from those data 
\citep[e.g.,][]{Deason_2011,Ness_2012,Ness_2013a,Ness_2013b,Casey_2012,Casey_2013,
Casey_2014a,Casey_2014b}.  % include references on: disk chemistry, halo structures
However, these quantities are not known for even 1\% of stars in the Milky Way.  
Stellar distances are famously imprecise \citep[e.g.,][]{van_Leeuwen_2007,
Jofre_2015,Madler_2016}, proper motions can be plagued by unquantified systematics
from the first epoch observations \citep[e.g.,][]{Casey_Schlaufman_2015}, and 
stellar spectroscopists frequently report significantly different chemical 
abundance patterns \citep{Smiljanic_2014}.  Countless long-standing scientific
questions could be definitively answered if these measurements were known with
high fidelity: \stub{x? \citep{someone}; y; and z, to list just a few}.

% Introducing Gaia
% Gaia will produce a very clear snapshot, producing astrometry space motions
% What will Gaia provide eventually.
% In the first data release what do we get.
% While this is an order of magnitude better than Hipparcos, if we want to use these data to understand the chemical evolution of the galaxy then we need additional information.
% until abundances are available from gaia RVS, we rely on ground-based spectroscopic surveys to provide the chemistry

The first \project{Gaia} data release will include positions, proper motions, and 
parallaxes for approximately two million stars in the Tycho-2 \citep{Hog_2000} 
catalog \citep{Michalik_2015a,Michalik_2015b}.  After cross-matching all major 
stellar spectroscopic surveys\footnote{Specifically we cross-matched the Tycho-2
catalog against the \project{APOGEE} \citep{Zasowski_2013}, \project{GALAH} 
\citep{DeSilva_2015}, \project{Gaia-ESO} \citep{Gilmore_2012,Randich_2013}, 
\project{LAMOST} \citep{Cui_2012}, and \project{RAVE} \citep{Steinmetz_2006} 
surveys}, we found the RAdial Velocity Experiment (hereafter \rave) survey to 
have the largest overlap with the first \project{Gaia} data release: 292,036 
stars.  We then used the \project{Gaia} universe model \citep{Robin_2012} to 
estimate the precision in parallax and proper motions that will be available in 
\project{Gaia} DR1 for every star in these cross-matches.  Comparing the expected
precision to what is currently available, we further found that the \project{RAVE}
survey will benefit most from \project{Gaia} DR1.  The distances of 63\% of the 
\project{RAVE}--\project{Gaia} DR1 overlap sample (182,862 stars) are expected to
improve with the first \project{Gaia} data release, and 47\% of stars likely to
have better proper motions (137,211 stars).  This motivated us to examine what 
chemical abundance information is currently available for \rave, and to evaluate
whether we could contribute to this existing set.

We briefly describe the \rave\ data in Section \ref{sec:data}, before explaining
the model we employ in Section \ref{sec:model}.  In Section \ref{sec:validation}
we outline a number of validation experiments, including: internal sanity checks,
comparisons with literature samples, and investigations to ensure our results
are consistent with expectations from astrophysics.  We provide concluding
remarks in Section \ref{sec:conclusion}, as well as a description of how to
access the data.

@@

The \rave\ survey started observations using the 6dF system on the 1.2~m UK 
Schmidt telescope at the Australian Astronomical Observatory\footnote{Formerly 
the Anglo-Australian Observatory}.  The large 5.7~degree field-of-view and 
robotic fibre positioner made for very efficient observing: up to 150 targets 
could be simultaneously obtained.  When observations concluded in April 2013, a 
total of 574,630 spectra had been collected of 483,330 unique objects.

RAVE is a magnitude-limited survey, targeting stars with apparent $I$-band
magnitudes between $9 < I < 13$.  The $I$ band was used as the selection
magnitude because it overlaps with the wavelength range that \rave\ operates in:
8410--8795~\AA.  This setup includes the \ion{Ca}{2} near-infrared triplet,
strong stellar absorption lines that are visible even in very low S/N data 
(i.e., they are very useful for radial velocity determination).

As the name implies, the \rave\ survey was principally designed to obtain radial
velocity measurements.  Detailed chemical abundances were a secondary science 
goal of the survey.  For this reason the distribution of S/N ratios for \rave\ 
spectra is considerably lower than stellar spectroscopic surveys where chemical
abundances are the defining objective.  The 
\rave\ spectra cover $\approx38$~nm in the infrared with an effective resolution
$\mathcal{R} \approx 7500$, and the S/N peaks at $\approx$50~per~pixel$^{-1}$.
For comparison, the \project{GALAH} survey \citep{DeSilva_2015} --- which was 
specifically constructed for detailed chemical abundance analyses --- includes a
wavelength range about 2.5 times larger at resolution $\mathcal{R} \approx 28,000$,
and target for S/N $\gtrsim100$.

Despite the relatively low resolution and S/N of the spectra, the \rave\ data
releases have demonstrated the team's enormous success in deriving radial 
velocities, stellar astrophysical parameters ($\teff$, $\logg$), as well as 
individual chemical abundances \citep{Steinmetz_2006,Zwitter_2008,Siebert_2011,
Kordopatis_2013, Kunder_2016}.  The data used for this analysis are the result
of a thorough reprocessing of all survey data.  This includes a detailed
re-reduction of the original data frames, an updated continuum-normalization
procedure, as well as an improvement on the quantification of flux errors in
each pixel.  At the end of this processing for each survey observation we are 
provided with: wavelengths, continuum-normalized fluxes, as well as 
$1\sigma$ uncertainties in flux values.  We refer the reader to the official 
fifth data release paper of the \rave\ survey, as presented by
\citet{Kunder_2016}, for more details of this reprocessing.

We chose not to re-normalize the spectra, as our tests proved to us that the
normalization procedure outlined in \citet{Kunder_2016} is sufficient for our
analysis.  We re-sampled all spectra onto a common rest-wavelength map.
Depending on the fibre used and stellar radial velocity, the rest-frame wavelength
range varied for each star.  Given that fluxes were unavailable in the edge
pixels for most stars, we excluded pixels outside of the rest wavelength range
$8423.2\,{\rm \AA} \le \lambda \le 8777.6\,{\rm \AA}$.

@@

% Describe The Cannon briefly
% training set from APOGEE

@@

% Initial tests using a single model for all RAVE spectra
% Censoring masks
% Priors on astrophysical parameters from isochrones

% Include velocities at test time?
% Train on fibre number? Fit for spectral resolution at test time?

@@

\acknowledgements
We thank Jonathan Bird (Vanderbilt), Sven Buder (MPIA), others??.
This research made use of: Astropy, a community-developed core Python package for
Astronomy \citep{astropy}, NASA's Astrophysics Data System Bibliographic Services;
and \project{TOPCAT} \citep{Taylor_2005}.
This work was partly supported by the European Union FP7 programme through ERC 
grant number 320360.
Funding for RAVE has been provided by: the Australian Astronomical Observatory; 
the Leibniz-Institut fuer Astrophysik Potsdam (AIP); the Australian National 
University; the Australian Research Council; the French National Research Agency;
the German Research Foundation (SPP 1177 and SFB 881); the European Research 
Council (ERC-StG 240271 Galactica); the Istituto Nazionale di Astrofisica at 
Padova; The Johns Hopkins University; the National Science Foundation of the USA
(AST-0908326); the W. M. Keck foundation; the Macquarie University; the 
Netherlands Research School for Astronomy; the Natural Sciences and Engineering 
Research Council of Canada; the Slovenian Research Agency; the Swiss National 
Science Foundation; the Science \& Technology Facilities Council of the UK; 
Opticon; Strasbourg Observatory; and the Universities of Groningen, Heidelberg 
and Sydney. The RAVE web site is https://www.rave-survey.org.

@@

>>>>

The Milky Way is considered to be the best laboratory for understanding galaxy
formation and evolution.  This premise hinges on the ability to precisely measure 
the position, motion, and chemistry of (many) individual stars, and use those data 
to infer the structure, kinematics, and chemical enrichment of the Galaxy 
\citep[e.g.,][]{Deason_2011,Ness_2012,Ness_2013a,Ness_2013b,Casey_2012,Casey_2013,
Casey_2014a,Casey_2014b}.  % include references on: disk chemistry, halo structures
However, these quantities are not known for even 1\% of stars in the Milky Way.  
Stellar distances are famously imprecise \citep[e.g.,][]{van_Leeuwen_2007,
Jofre_2015,Madler_2016}, proper motions can be plagued by unquantified systematics
from the first epoch observations \citep[e.g.,][]{Casey_Schlaufman_2015}, and 
stellar spectroscopists frequently report significantly different chemical 
abundance patterns \citep{Smiljanic_2014}.  Countless long-standing astrophysics
problems could be resolved if those measurements were known with high fidelity: 
\stub{x? \citep{someone}; y; and z, to list just a few}.

\stub{Introducing Gaia; rely on ground-based spectroscopic surveys for chemistry}
% Introducing Gaia
% Gaia will produce a very clear snapshot, producing astrometry space motions
% What will Gaia provide eventually.
% In the first data release what do we get.
% While this is an order of magnitude better than Hipparcos, if we want to use these data to understand the chemical evolution of the galaxy then we need additional information.
% until abundances are available from gaia RVS, we rely on ground-based spectroscopic surveys to provide the chemistry

The first \project{Gaia} data release will include positions, proper motions, and 
parallaxes for approximately two million stars in the Tycho-2 \citep{Hog_2000} 
catalog \citep{Michalik_2015a,Michalik_2015b}.  After cross-matching all major 
stellar spectroscopic surveys\footnote{Specifically we cross-matched the Tycho-2
catalog against the \project{APOGEE} \citep{Zasowski_2013}, \project{GALAH} 
\citep{DeSilva_2015}, \project{Gaia-ESO} \citep{Gilmore_2012,Randich_2013}, 
\project{LAMOST} \citep{Cui_2012}, and \project{RAVE} \citep{Steinmetz_2006} 
surveys}, we found the RAdial Velocity Experiment (hereafter \rave) survey to 
have the largest overlap with the first \project{Gaia} data release: 292,036 
stars.  We then used the \project{Gaia} universe model \citep{Robin_2012} to 
estimate the precision in parallax and proper motions that will be available in 
\project{Gaia} DR1 for every star in these cross-matches.  Comparing the expected
precision to what is currently available, we further found that the \project{RAVE}
survey will benefit most from \project{Gaia} DR1.  The distances of 63\% of the 
\project{RAVE}--\project{Gaia} DR1 overlap sample (182,862 stars) are expected to
improve with the first \project{Gaia} data release, and 47\% of stars likely to
have better proper motions (137,211 stars).  This motivated us to examine what 
chemical abundance information was available for \rave, and to evaluate whether 
we could enable new chemodynamic studies by contributing to this existing set of
chemical abundances.

We briefly describe the \rave\ data in section \ref{sec:data}, before explaining
the model we employ in section \ref{sec:model}.  In section \ref{sec:validation}
we outline a number of validation experiments, including: internal sanity checks,
comparisons with literature samples, and investigations to ensure our results
are consistent with expectations from astrophysics.  We provide concluding
remarks in section \ref{sec:conclusion}, as well as a description of how to
access our results electronically.

@@

\rave\ is a magnitude-limited stellar spectroscopic survey of the (nearby) Milky Way,
principally designed to measure radial velocities for up to $10^6$ stars.
Observations were conducted on the 1.2~m UK Schmidt telescope at the Australian 
Astronomical Observatory\footnote{Formerly the Anglo-Australian Observatory} from 
2003--2013.  The large 5.7~degree field-of-view and robotic fibre positioner (6dF)
made for very efficient observing: spectra for up to 150 targets could be 
simultaneously acquired.  When observations concluded in April 2013, a total of 
574,630 spectra had been collected of 483,330 unique objects.

The target selection for \rave\ is solely based on the $I$-band apparent magnitude: 
$9 < I < 13$.  The $I$ band was used because it overlaps with the wavelength range 
that \rave\ operates in:  8410--8795~\AA.  This setup includes the \ion{Ca}{2} 
near-infrared triplet.  These stellar absorption lines are dominated by pressure 
broadening, which makes them strong and therefore visible even in very low S/N data
or metal-poor stars.  This spectral region is also well-studied.  It is one of the
key setups used for the ground-based high-resolution \project{Gaia-ESO} survey
\citep{Gilmore_2012,Randich_2013}, and the \rave\ resolution and wavelength
coverage is comparable to the Radial Velocity Spectrometer on board the \project{Gaia}
space telescope \citep{Recio-Blanco_2016}.

The exposure times (and resultant S/N ratios) for \rave\ were optimised to obtain
radial velocities for as many stars as possible.  Detailed chemical abundances were
always an important science goal of the survey, but this was a secondary objective.  
For this reason the distribution of S/N ratios of \rave\ spectra is considerably 
lower than other stellar spectroscopic surveys where chemical abundances are the 
primary motivation.  The \rave\ spectra have an effective resolution 
$\mathcal{R} \approx 7500$ and the distribution of S/N ratios peaks at 
$\approx$50~per~pixel$^{-1}$.  For comparison, the \project{GALAH} survey 
\citep{DeSilva_2015} --- which was specifically constructed for detailed chemical 
abundance analyses --- includes a wavelength range about 2.5 times larger at 
resolution $\mathcal{R} \approx 28,000$, and yet the \project{GALAH} project still 
targets for S/N $\gtrsim100$.

Despite the relatively low resolution and S/N of the spectra compared to other
surveys, the \rave\ data releases have demonstrated the team's enormous success 
in deriving radial velocities, stellar astrophysical parameters ($\teff$, $\logg$),
as well as individual chemical abundances \citep{Steinmetz_2006,Zwitter_2008,
Siebert_2011,Kordopatis_2013, Kunder_2016}.  In this work we make use of spectra
that has been reprocessed for the fifth \rave\ data release.  These re-processing
steps include: a detailed re-reduction of all the original data frames, with flux
variances propagated at every step; an updated continuum-normalization procedure;
as well as revised determinations of stellar radial velocities and morphological
classifications. At the end of this processing for each survey observation we are 
provided with: rest-frame wavelengths, continuum-normalized fluxes, fractional 
uncertainties in continuum-normalized flux values, as well as relevant metadata.  
We refer the reader to the official fifth data release paper of the \rave\ survey, 
as presented by \citet{Kunder_2016}, for more details of this re-processing.

Given the high-quality of the normalization performed by the \rave\ team, we chose
not to re-normalize the spectra.  Our tests demonstrated that the procedure 
outlined in \citet{Kunder_2016} is sufficient for our analysis procedure. Therefore,
there is a limited number of pre-processing steps that we conducted before our
analysis could begin.  Firstly, we calculated inverse variance arrays from the
fractional uncertainties provided, then re-sampled the flux and inverse variance
arrays onto a common rest-wavelength map for all stars.  Depending on the fibre 
used and the stellar radial velocity, the range of rest-frame wavelength values
varied for each star.  Given that fluxes were unavailable in the edge pixels for 
most stars, we excluded pixels outside of the rest wavelength range 
$8423.2\,{\rm \AA} \le \lambda \le 8777.6\,{\rm \AA}$.

@@

\stub{data-driven goodness}
% Describe The Cannon briefly, refer to Ness.

\subsection{The training set}
\label{sec:the-training-set}

% training set from APOGEE or K2/EPIC?!!

@@

% Censoring masks on abundance labels
% Priors on astrophysical parameters from isochrones

% Include velocities at test time?

@@

\acknowledgements
A.~R.~C. warmly thanks Jonathan Bird (Vanderbilt) and Sven Buder (MPIA).
This research made use of: Astropy, a community-developed core Python package for
Astronomy \citep{astropy}, NASA's Astrophysics Data System Bibliographic Services;
and \project{TOPCAT} \citep{Taylor_2005}.
This work was partly supported by the European Union FP7 programme through ERC 
grant number 320360.
Funding for RAVE has been provided by: the Australian Astronomical Observatory; 
the Leibniz-Institut fuer Astrophysik Potsdam (AIP); the Australian National 
University; the Australian Research Council; the French National Research Agency;
the German Research Foundation (SPP 1177 and SFB 881); the European Research 
Council (ERC-StG 240271 Galactica); the Istituto Nazionale di Astrofisica at 
Padova; The Johns Hopkins University; the National Science Foundation of the USA
(AST-0908326); the W. M. Keck foundation; the Macquarie University; the 
Netherlands Research School for Astronomy; the Natural Sciences and Engineering 
Research Council of Canada; the Slovenian Research Agency; the Swiss National 
Science Foundation; the Science \& Technology Facilities Council of the UK; 
Opticon; Strasbourg Observatory; and the Universities of Groningen, Heidelberg 
and Sydney. The RAVE web site is https://www.rave-survey.org.

@@

\bibitem[Recio-Blanco et al.(2016)]{Recio-Blanco_2016} Recio-Blanco, A., de Laverny, P., Allende Prieto, C., et al.\ 2016, \aap, 585, A93
